full title

### 2020-09-11 paper/pldi21/formal.tex
[PLDI21] Lazy Concrete Interpretation / Combined Analysis

We define a \textit{memory abstraction} $\powerset{\memset}
\galois{\alpha_\memset}{\gamma_\memset} \absmemset$ as a Galois connection
between sets of concrete memories and abstract memories.  An abstract memory
$\absmem \in \absmemset: \abslocset \finmap \absvalset$ is a finite mapping from
abstract locations to abstract values.  The \textit{strong update} in memory
abstraction is important to increase the precision of static analysis.  For
sound analysis, a memory update for a specific abstract location should join old
abstract values with new abstract values, which calls a \textit{weak update}.
It degrades the analysis precision because of the sprious values.  To overcome
such degradation, researchers proposed the abstract
counting~\cite{abstract-gc-counting, revisit-recency} to track how many times an
abstract location has been allocated and to apply strong updates to singleton
abstract locations.

We extended abstract states with an abstract counting operator $\abscount{-}:
\abslocset \rightarrow \abscountset = \{ \abszero, \absone, \absmany \}$, which
is a function that takes an abstract location and returns its abstract count;
$\abszero$ denotes that it have never been allocated, $\absone$ once, and
$\absmany$ more than equals to twice.

\subsection{On-Demand Access Aanlysis}

>>>>

We define a \textit{memory abstraction} $\powerset{\memset}
\galois{\alpha_\memset}{\gamma_\memset} \absmemset$ as a Galois connection
between sets of concrete memories and abstract memories.  An abstract memory
$\absmem \in \absmemset: \abslocset \finmap \absvalset$ is a finite mapping from
abstract locations to abstract values.  The \textit{strong update} in memory
abstraction is important to increase the precision of static analysis.  For
sound analysis, a memory update for a specific abstract location should join old
abstract values with new abstract values, called as a \textit{weak update}.
It degrades the analysis precision because of the sprious values.  To overcome
such degradation, researchers proposed the abstract
counting~\cite{abstract-gc-counting, revisit-recency} to track how many times an
abstract location has been allocated and to apply strong updates to singleton
abstract locations.

We extend abstract states with an abstract counting operator $\abscount{-}:
\abslocset \rightarrow \abscountset = \{ \abszero, \absone, \absmany \}$, which
is a function that takes an abstract location and returns its abstract count;
$\abszero$ denotes that it have never been allocated, $\absone$ once, and
$\absmany$ more than or equals to twice.

\subsection{Lazy Concrete Interpretation}

We define a \textit{lazy concrete interpretation} by extending the transition
relation $\trans$ as a symbolic transition relation $\symbtrans$ on abstract
states. First, we extends the concrete states $\stset$ to \textit{symbolic
states} $\symbstset = (\labset \times \symbmemset) \cup \{ \excst \}$ with
symbolic memories $\symbmemset$ consisting of symbolic values $\symb \in
\symbset$ and a special exception state $\excst$. Each symbolic value denotes an
abstract value cannot be concretized to a singleton value.  A symbolic state
\textit{alwyas} has a symbolic transition relation with a single symbolic state.
When a symbolic state has a non-determinism or accesses symbolic values, it has
the relation with the exception state: $\symbst \symbtrans \excst$.  Thus, the lazy
concrete interpretation does not care about the non-deterministic semantics or
symbolic values.

We define a Galois connection $\symbstset
\galois{\alpha_\symbset}{\gamma_\symbset} \absdom$ to freely convert an abstract
state to the corresponding symbolic state and vice versa.  In the concretization
function $\gamma_\symbset$, if an abstract value $\absval$ represents a
singleton value, the function transforms the abstract value to the
corresponding concrete value.  Otherwise, it keeps the abstract value as a
symbolic value in the result symbolic state.

\subsection{Combined Analysis}

We define the \textit{combined analysis} of the abstract interpretation and the
lazy concrete interpretation by extending the view transition as follows:
\[
  \viewtrans{\view'}{\view}_\symbset(\abselem) = \left\{
    \begin{array}{ll}
      \abselem_0 & \text{if} \; \symbst' \symbtrans \excst\\
      \alpha_\symbset(\symbst) & \text{if} \; \symbst' \symbtrans \symbst
      \wedge\\
      & \phantom{\text{if}} \; \gamma \circ \alpha_\symbset(\symbst)
      \cap \viewmap(\view) \neq \varnothing\\ \bot & \text{otherwise}\\
    \end{array}
  \right.
\]
where $\symbst' = \gamma_\symbset \circ \abselem(\view')$ and $\abselem_0 =
\viewtrans{\view'}{\view}(\abselem)$.  The combined analysis first tries to
convert the current abstract state $\abselem(\view')$ to the corresponding
symbolic state $\symbst'$ and performs the lazy concrete evaluation.  When the
evaluation produces the exception state $\excst$, it utilizes the abstract
interpretation instead of the lazy concrete interpretation.  Otherwise, the lazy
concrete interpretation successfully produces a new symbolic state $\symbst$.
If the symbolic state $\symbst$ is related to the target view $\view$, the
result is the corresponding abstract state $\alpha_\symbset(\symbst)$ of the
symbolic state $\symbst$. Otherwise, the result becomes the bottom $\bot$.
